---
title: "Automatic walking strides segmentation from  wrist-worn sensor accelerometry data collected in the free-living"
output: 
  github_document:
    toc: true
    toc_depth: 4
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  message = FALSE,
  warning = FALSE,
  # cache.lazy = FALSE,
  comment = "#>"
)
```

Below, we demonstrate our method for automatic walking strides segmentation from wrist-worn sensor accelerometry data collected in the free-living environment. 

This repository accompanies "*Estimation of free-living walking cadence from wrist-worn sensor accelerometry data and its association with SF-36 quality of life scores*" [manuscript][1] in which the method is proposed. The method is implemented as the `segmentWalking()` function in `adept` R package ([CRAN index](https://cran.r-project.org/web/packages/adept/index.html)). Below, we demonstrate the method with the use of independent data sample (not used in the manuscript study). 

### Accelerometry data sample 

#### Description 

The directory `data/` contains data files: 

* `acc_raw_left_wrist.rds`
* `acc_raw_right_wrist.rds`

which are a sample of raw accelerometry data collected consecutively for 2 days, starting and ending at midnight of a week day. Data were collected at the sampling frequency of 100 Hz with two ActiGraph GT9X Link sensors located at the left wrist and right wrist, respectively. One data observation includes `(x,y,z)` values collected from each of the three axes at the same time. At the sampling frequency of 100 Hz, there were 100 observations collected per second (17,280,000 observations per 2 days) for each of the two sensors.

Below, we read the data and plot a few first and last observations from left wrist data set. 

```{r read_data, message = FALSE}
library(tidyverse)
library(lubridate)
library(adeptdata) # for stride pattern templates data 
library(adept)

# read raw accelerometry data from left wrist, right wrist
dat_lw <- readRDS(paste0(here::here(), "/data/acc_raw_left_wrist.rds")) %>% as_tibble()
dat_rw <- readRDS(paste0(here::here(), "/data/acc_raw_right_wrist.rds")) %>% as_tibble()

# define data collection POSIXct (used *fake* day date of Feb 1, 2021)
obs_datetime_seq <- seq(ymd_hms("2021-02-01 00:00:00"), by = 0.01, length.out = nrow(dat_lw))
dat_lw <- mutate(dat_lw, obs_datetime = obs_datetime_seq, .before = "x")
dat_rw <- mutate(dat_rw, obs_datetime = obs_datetime_seq, .before = "x")
```

```{r}
rbind(head(dat_lw, 3), tail(dat_lw, 3))
```

Data were collected from a healthy 28F person.

<details>
  <summary>(Click to see IRB note.)</summary>
  
The raw accelerometry data in `data/` package were collected from sensors worn by Marta Karas, an author of this repository. The IRB Office Determination Request Form for Primary (New) Data Collection request form was submitted in regard to the collection and further publishing of these data. Based on preliminary review of the request form submitted, it was determined that the data collection and further data publishing activity described in the determination request does not qualify as human subjects research as defined by DHHS regulations 45 CFR 46.102, and does not require IRB oversight.
</details>

#### Raw data visualization  

One way to visualize raw accelerometry data is to plot it as a three-dimensional time-series `(x,y,z)`. Here, we plot data from three different time frames, each of 4 seconds length, simultaneously for data collected at left wrist and right wrist. 

<details> <summary>(Click to see the code.)</summary>
```{r plot_xyz, fig.width=7.5, fig.height=4.5, eval = FALSE}
# define time frame start values for data subset
t1 <- ymd_hms("2021-02-01 09:57:11") 
t2 <- ymd_hms("2021-02-01 09:59:41") 
t3 <- ymd_hms("2021-02-01 10:03:47") 

# combine data from two sensors, subset to keep only selected time windows 
dat <- rbind(dat_lw, dat_rw)
dat_sub <- dat %>%
  filter((obs_datetime >= t1 & obs_datetime < t1 + as.period(4, "seconds")) | 
           (obs_datetime >= t2 & obs_datetime < t2 + as.period(4, "seconds")) | 
           (obs_datetime >= t3 & obs_datetime < t3 + as.period(4, "seconds")) ) %>%
  mutate(dt_floor = paste0("time frame start: ", 
                           floor_date(obs_datetime, unit = "minutes")))

# plot (x,y,z) values 
dat_sub %>%
  select(-obs_idx) %>% 
  pivot_longer(cols = -c(dt_floor, obs_datetime, loc_id)) %>%
  ggplot(aes(x = obs_datetime, y = value, color = name)) + 
  geom_line(size = 0.3) + 
  facet_grid(loc_id ~ dt_floor, scales = "free_x") + 
  theme_minimal(base_size = 10) + 
  labs(x = "Time [s]", 
       y = "Acceleration [g]", 
       color = "Accelerometer axis of measurement: ") + 
  theme(legend.position = "top")
```
</details>

```{r plot_xyz_2, fig.width=7.5, fig.height=4.5, echo = FALSE}
# define time frame start values for data subset
t1 <- ymd_hms("2021-02-01 09:57:11") 
t2 <- ymd_hms("2021-02-01 09:59:41") 
t3 <- ymd_hms("2021-02-01 10:03:47") 

# combine data from two sensors, subset to keep only selected time windows 
dat <- rbind(dat_lw, dat_rw)
dat_sub <- dat %>%
  filter((obs_datetime >= t1 & obs_datetime < t1 + as.period(4, "seconds")) | 
           (obs_datetime >= t2 & obs_datetime < t2 + as.period(4, "seconds")) | 
           (obs_datetime >= t3 & obs_datetime < t3 + as.period(4, "seconds")) ) %>%
  mutate(dt_floor = paste0("time frame start: ", 
                           floor_date(obs_datetime, unit = "minutes")))

# plot (x,y,z) values 
dat_sub %>%
  select(-obs_idx) %>% 
  pivot_longer(cols = -c(dt_floor, obs_datetime, loc_id)) %>%
  ggplot(aes(x = obs_datetime, y = value, color = name)) + 
  geom_line(size = 0.3) + 
  facet_grid(loc_id ~ dt_floor, scales = "free_x") + 
  theme_minimal(base_size = 10) + 
  labs(x = "Time [s]", 
       y = "Acceleration [g]", 
       color = "Accelerometer axis of measurement: ") + 
  theme(legend.position = "top")
```

Vector magnitude (VM) is often used to reduce the dimensionality of accelerometry time-series `(x,y,z)`. VM is computed as `vm = sqrt(x^2 + y^2 + z^2)` at each time point resulting in 1- instead of 3-dimensional time-series.

<details> <summary>(Click to see the code.)</summary>
```{r plot_vm, fig.width=7.5, fig.height=4.1, eval = FALSE}
# plot vector magnitude values 
dat_sub %>%
  mutate(vm = sqrt(x^2 + y^2 + z^2)) %>%
  ggplot(aes(x = obs_datetime, y = vm)) + 
  geom_line(size = 0.3) + 
  facet_grid(loc_id ~ dt_floor, scales = "free_x") + 
  theme_minimal(base_size = 10) + 
  labs(x = "Time [s]", y = "VM") 
```
</details>

```{r plot_vm_2, fig.width=7.5, fig.height=4.1, echo = FALSE}
# plot vector magnitude values 
dat_sub %>%
  mutate(vm = sqrt(x^2 + y^2 + z^2)) %>%
  ggplot(aes(x = obs_datetime, y = vm)) + 
  geom_line(size = 0.3) + 
  facet_grid(loc_id ~ dt_floor, scales = "free_x") + 
  theme_minimal(base_size = 10) + 
  labs(x = "Time [s]", y = "VM") 
```

```{r plot_vmc, fig.width=7.5, fig.height=4.1, eval = FALSE, echo = FALSE}
# <details> <summary>(Click to see the code.)</summary>

# In reality, it is often challenging to make a plot of all data points collected at sampling frequency of 100 Hz even in a few minutes-long time-series. A way to summarize accelerometry data of such high density is to use vector magnitude count (VMC; also known as the mean amplitude deviation, MAD). VMC over time window of length H is defined as mean of absolute value of `vm` deviations from its mean in the window. 
# 
# Here, we compute VMC over 10 minutes-long time window.

# function to compute VMC in a VM window
vmc_func <- function(vm_win){
  mean(abs(vm_win - mean(vm_win)))
}

# compute VMC vector in 10-minutes windows
vm <- sqrt(dat$x^2 + dat$y^2 + dat$z^2)
win_vl <- 100 * 60 * 10
rn_seq <- seq(from = 1, to = nrow(dat), by = win_vl)
vmc_vec <- sapply(rn_seq, function(rn_i){
  vm_win_idx <- rn_i : (rn_i + win_vl - 1)
  vm_win <- vm[vm_win_idx]
  vmc_func(vm_win)
})

# plot VMC
plt_df <- dat[rn_seq, ] %>% select(obs_datetime, loc_id) %>%
  mutate(vmc = vmc_vec)
ggplot(plt_df, aes(x = obs_datetime, y = vmc)) +
  facet_grid(loc_id ~ .) +
  geom_line(size = 0.3) +
  theme_minimal(base_size = 10) +
  labs(x = "Time", y = "VMC")
```

### Segment individual walking strides 

#### Algorithm

We use `segmentWalking()` function in `adept` R package to automatically segment individual walking strides from the raw accelerometry data collected with a wrist-worn sensor. 

First, the function first uses Adaptive Empirical Pattern Transformation (ADEPT) method (see [ADEPT manuscript][2]) to segment walking stride patterns. In short, ADEPT uses a predefined template and detects its repetitions in data by maximizing the local correlation between scale-transformed template's versions and the observed data at every time point. Multiple baseline templates can be used simultaneously to best fit various shapes of strides. The ADEPT result describes each segmented stride by its: 

- (a) start time, 
- (b) duration (seconds),  
- (c) correlation with best-fit template.

Second, the function filters the patterns segmented from the data to keep those which plausibly and with high specificity correspond to walking activity. Methodological details are reported in the method [manuscript][1].

#### Stride pattern templates

Below, we use three wrist-specific stride templates publicly available in [adeptdata](https://cran.r-project.org/web/packages/adeptdata/index.html) R package. 

<details> <summary>(Click to see the code.)</summary>
```{r plot_templates, fig.width=7.5, fig.height=3, eval = FALSE}
# pull 3 x 200 matrix with 3 distinct stride pattern templates
# (attached in adeptdata R package)
templates_mat <- stride_template$left_wrist[[3]]

plt_df <- 
  templates_mat %>% 
  as.data.frame() %>%
  rename_all(~as.character(seq(0, 1, length.out = ncol(templates_mat)))) %>%
  mutate(template_idx = row_number()) %>%
  pivot_longer(cols = -template_idx) %>%
  mutate(name = as.numeric(name),
         template_idx = factor(template_idx, labels = paste0("Template ", 1 : 3)))

ggplot(plt_df, aes(x = name, y = value)) + 
  geom_line() + 
  facet_wrap(~ template_idx) +
  theme_minimal(base_size = 10) +
  labs(x = "Templaste phase", y = "VM") + 
  labs(x = "Time [s]", y = "VM") 
```
</details>

```{r plot_templates_2, fig.width=7.5, fig.height=3, echo = FALSE}
# pull 3 x 200 matrix with 3 distinct stride pattern templates
# (attached in adeptdata R package)
templates_mat <- stride_template$left_wrist[[3]]

plt_df <- 
  templates_mat %>% 
  as.data.frame() %>%
  rename_all(~as.character(seq(0, 1, length.out = ncol(templates_mat)))) %>%
  mutate(template_idx = row_number()) %>%
  pivot_longer(cols = -template_idx) %>%
  mutate(name = as.numeric(name),
         template_idx = factor(template_idx, labels = paste0("Template ", 1 : 3)))

ggplot(plt_df, aes(x = name, y = value)) + 
  geom_line() + 
  facet_wrap(~ template_idx) +
  theme_minimal(base_size = 10) +
  labs(x = "Templaste phase", y = "VM") + 
  labs(x = "Time [s]", y = "VM") 
```


#### Segmentation

We use `segmentWalking()` method from `adept` R package to segment walking separately from data from (a) left wrist, (b) right-wrist worn sensor. Default algorithm parameters are optimized for a wrist-worn sensor. 

<details> <summary>(Click to see the code -- part 1.)</summary>
```{r, eval = FALSE}
# convert templates matrix to templates list 
templates_list <-  list(templates_mat[1, ], templates_mat[2, ], templates_mat[3, ])

# segment walking from data collected with a left wrist-worn sensor
t1 <- Sys.time()
out_lw <- segmentWalking(
  xyz = dat_lw %>% select(all_of(c("x", "y", "z"))),
  xyz.fs = 100,
  template = templates_list,
  run.parallel = TRUE,
  run.parallel.cores = 8)
Sys.time() - t1
# Time difference of 10.45637 mins

# segment walking from data collected with a right wrist-worn sensor
t1 <- Sys.time()
out_rw <- segmentWalking(
  xyz = dat_rw %>% select(all_of(c("x", "y", "z"))),
  xyz.fs = 100,
  template = templates_list,
  run.parallel = TRUE,
  run.parallel.cores = 8)
Sys.time() - t1
# Time difference of 5.958216 mins

# save precomputed results for faster README compilation 
saveRDS(out_lw, paste0(here::here(), "/data/out_lw.rds"))
saveRDS(out_rw, paste0(here::here(), "/data/out_rw.rds"))
```
</details>

<details> <summary>(Click to see the code -- part 2.)</summary>
```{r read_precomputed}
# read precomputed results for faster README compilation 
out_lw <- readRDS(paste0(here::here(), "/data/out_lw.rds"))
out_rw <- readRDS(paste0(here::here(), "/data/out_rw.rds")) 

# filter output to keep rows corresponding to identified walking
out_lw_1 <- out_lw %>% filter(is_walking_i == 1)
out_rw_1 <- out_rw %>% filter(is_walking_i == 1)

head(out_lw_1)
```
</details>

#### Visualization of segmented walking strides

Below, we visualize first K=50 strides segmented from (a) left wrist, (b) right-wrist worn sensor.  

<details> <summary>(Click to see the code.)</summary>
```{r vizu_segmented, fig.width=7.5, fig.height=3, eval = FALSE}
# get VM from raw data frames
dat_lw_vm <- sqrt(dat_lw$x^2 + dat_lw$y^2 + dat_lw$z^2)
dat_rw_vm <- sqrt(dat_rw$x^2 + dat_rw$y^2 + dat_rw$z^2)

K <- 50
T_i_max <- max(c(out_lw_1$T_i, out_rw_1$T_i))
stride_dat <- numeric()
stride_id  <- numeric()
stride_loc <- numeric()
for (k in 1 : K){ # k <- 1
  # left wrist data 
  dat_idx_k  <- (out_lw_1[k, "tau_i"]) : (out_lw_1[k, "tau_i"] + out_lw_1[k, "T_i"] - 1)
  dat_k      <- dat_lw_vm[dat_idx_k]
  stride_dat <- c(stride_dat, dat_k)
  stride_id  <- c(stride_id, rep(k, length(dat_k)))
  stride_loc <- c(stride_loc, rep("left_wrist", length(dat_k)))
  # right wrist data 
  dat_idx_k  <- (out_rw_1[k, "tau_i"]) : (out_rw_1[k, "tau_i"] + out_rw_1[k, "T_i"] - 1)
  dat_k      <- dat_rw_vm[dat_idx_k]
  stride_dat <- c(stride_dat, dat_k)
  stride_id  <- c(stride_id, rep(k, length(dat_k)))
  stride_loc <- c(stride_loc, rep("right_wrist", length(dat_k)))
}

# define plot data frame
plt_df <- 
  data.frame(stride_dat, stride_id, stride_loc) %>%
  group_by(stride_id, stride_loc) %>%
  mutate(stride_duration = row_number() - 1,
         stride_duration = stride_duration / 100) %>%
  ungroup()
# point-wise aggregate of strides
plt_df_agg <- 
  plt_df %>% 
  group_by(stride_loc, stride_duration) %>% 
  summarise(stride_dat = mean(stride_dat))  %>%
  ungroup()

ggplot(plt_df, aes(x = stride_duration, y = stride_dat, group = stride_id)) + 
  geom_line(alpha = 0.2) +
  facet_wrap(~ stride_loc) +
  theme_minimal(base_size = 10) +
  labs(x = "Stride duration [s]", y = "VM") 
```
</details>

```{r vizu_segmented_2, fig.width=7.5, fig.height=3, echo = FALSE}
# get VM from raw data frames
dat_lw_vm <- sqrt(dat_lw$x^2 + dat_lw$y^2 + dat_lw$z^2)
dat_rw_vm <- sqrt(dat_rw$x^2 + dat_rw$y^2 + dat_rw$z^2)

K <- 50
T_i_max <- max(c(out_lw_1$T_i, out_rw_1$T_i))
stride_dat <- numeric()
stride_id  <- numeric()
stride_loc <- numeric()
for (k in 1 : K){ # k <- 1
  # left wrist data 
  dat_idx_k  <- (out_lw_1[k, "tau_i"]) : (out_lw_1[k, "tau_i"] + out_lw_1[k, "T_i"] - 1)
  dat_k      <- dat_lw_vm[dat_idx_k]
  stride_dat <- c(stride_dat, dat_k)
  stride_id  <- c(stride_id, rep(k, length(dat_k)))
  stride_loc <- c(stride_loc, rep("left_wrist", length(dat_k)))
  # right wrist data 
  dat_idx_k  <- (out_rw_1[k, "tau_i"]) : (out_rw_1[k, "tau_i"] + out_rw_1[k, "T_i"] - 1)
  dat_k      <- dat_rw_vm[dat_idx_k]
  stride_dat <- c(stride_dat, dat_k)
  stride_id  <- c(stride_id, rep(k, length(dat_k)))
  stride_loc <- c(stride_loc, rep("right_wrist", length(dat_k)))
}

# define plot data frame
plt_df <- 
  data.frame(stride_dat, stride_id, stride_loc) %>%
  group_by(stride_id, stride_loc) %>%
  mutate(stride_duration = row_number() - 1,
         stride_duration = stride_duration / 100) %>%
  ungroup()
# point-wise aggregate of strides
plt_df_agg <- 
  plt_df %>% 
  group_by(stride_loc, stride_duration) %>% 
  summarise(stride_dat = mean(stride_dat))  %>%
  ungroup()

ggplot(plt_df, aes(x = stride_duration, y = stride_dat, group = stride_id)) + 
  geom_line(alpha = 0.2) +
  facet_wrap(~ stride_loc) +
  theme_minimal(base_size = 10) +
  labs(x = "Stride duration [s]", y = "VM") 
```


### Estimation of walking cadence (steps/s)

#### Daily walking cadence 

The data used in this tutorial consists of two days of continuous monitoring. We use 

segmented strides to estimate the walking cadence (number of steps per second) at every time of the day when walking was identified. Next, we compute the mode of all estimated cadences during the day was computed -- a *daily cadence* measurement. 

<details> <summary>(Click to see the code.)</summary>
```{r vizu_daily_candece, fig.width=7.5, fig.height=3, eval = FALSE}
# function to compute mode
estimate_mode <- function(x) {
  d <- density(x)
  d$x[which.max(d$y)]
}

# data collection frequency
hz_val <- 100

# combine result from walking segmentation from left and right wrist  
out_comb_1 <- rbind(
  out_lw_1 %>% mutate(loc_id = "left_wrist"),
  out_rw_1 %>% mutate(loc_id = "right_wrist")
) %>% 
  mutate(
    # define stride duration (in seconds) as number of observations in stride 
    # (stride := two subsequent steps)
    # over data frequency (number of observations per 1 second)
    dur_i = T_i / hz_val, 
    # define cadence as number of steps per 1 second
    cadence_i = 2/dur_i,
    # define collection day "ID" 
    day_id = ceiling(tau_i / (hz_val * 60 * 60 * 24)),
    day_id = factor(day_id, levels = c(1,2), labels = paste0("Day ", 1:2))
  ) 

# aggregate the cadence measurements by computing empirical mode
# (1 value per day)
out_comb_1_mode <- 
  out_comb_1 %>%
  group_by(loc_id, day_id) %>%
  summarise(stride_duration_mode = estimate_mode(dur_i),
            cadence_mode = estimate_mode(cadence_i)) %>%
  arrange(day_id, loc_id) %>%
  as.data.frame()

ggplot(out_comb_1, aes(x = cadence_i)) + 
  geom_density() + 
  facet_grid(loc_id ~ day_id) + 
  geom_vline(data = out_comb_1_mode, aes(xintercept = cadence_mode), 
             linetype = 2, color = "blue") +
  theme_minimal(base_size = 10) + 
  labs(x = "Cadence [steps/s]", title = "Vertical dashed line: empirical mode")
```
</details>

```{r vizu_daily_candece_2, fig.width=7.5, fig.height=4, echo = FALSE}
# function to compute mode
estimate_mode <- function(x) {
  d <- density(x)
  d$x[which.max(d$y)]
}

hz_val <- 100
out_comb_1 <- rbind(
  out_lw_1 %>% mutate(loc_id = "left_wrist"),
  out_rw_1 %>% mutate(loc_id = "right_wrist")
) %>% 
  mutate(dur_i = T_i / hz_val, 
         cadence_i = 2/dur_i,
         day_id = ceiling(tau_i / (hz_val * 60 * 60 * 24)),
         day_id = factor(day_id, levels = c(1,2), labels = paste0("Day ", 1:2))) 
out_comb_1_mode <- 
  out_comb_1 %>%
  group_by(loc_id, day_id) %>%
  summarise(stride_duration_mode = estimate_mode(dur_i),
            cadence_mode = estimate_mode(cadence_i)) %>%
  arrange(day_id, loc_id) %>%
  as.data.frame()

ggplot(out_comb_1, aes(x = cadence_i)) + 
  geom_density() + 
  facet_grid(loc_id ~ day_id) + 
  geom_vline(data = out_comb_1_mode, aes(xintercept = cadence_mode), 
             linetype = 2, color = "blue") +
  theme_minimal(base_size = 10) + 
  labs(x = "Cadence [steps/s]", title = "Vertical dashed line: daily cadence (empirical mode)")
```

The table summarizes empirical mode of stride duration [s] and cadence [steps/s]: 

<details> <summary>(Click to see the code.)</summary>
```{r vizu_daily_candeceB}
print(out_comb_1_mode)
```
</details>

```{r vizu_daily_candeceB_2, echo = FALSE}
print(out_comb_1_mode)
```



#### Hourly walking cadence 

Similarly, next, we compute the mode of all estimated cadences during the day-hour was computed. We visualize hourly cadence estimated for day-hours where the number of estimated walking strides is more or equal 10. 

<details> <summary>(Click to see the code.)</summary>
```{r vizu_daily_candece_hourly, fig.width=7.5, fig.height=3, eval = FALSE}
# compute hour ID  
out_comb_1 <- 
  out_comb_1 %>%
  mutate(
    hour_id = floor(tau_i / (hz_val * 60 * 60)),
    hour_id = hour_id %% 24
  )

# aggregate the cadence measurements by computing empirical mode
# (1 value per day-hour)
out_comb_1_mode_hourly <- 
  out_comb_1 %>%
  group_by(loc_id, day_id, hour_id) %>%
  summarise(cadence_mode = estimate_mode(cadence_i),
            cnt = n()) %>%
  as.data.frame()

ggplot(out_comb_1_mode_hourly %>% filter(cnt > 10), 
       aes(x = hour_id, y = cadence_mode, group = 1)) + 
  geom_line(linetype = 2, size = 0.5, color = "grey") + 
  geom_point(aes(size = cnt, alpha = cnt)) + 
  facet_grid(loc_id ~ day_id) + 
  theme_minimal(base_size = 10) + 
  labs(x = "Hour of a day", 
       y = "Cadence [steps/s]",
       alpha = "Segmented\nstrides\ncount",
       size = "Segmented\nstrides\ncount") + 
  scale_x_continuous(limits = c(0, 23), breaks = seq(0, 23, by = 2))
```
</details>

```{r vizu_daily_candece_hourly_2, fig.width=7.5, fig.height=4, echo = FALSE}
# compute hour ID  
out_comb_1 <- 
  out_comb_1 %>%
  mutate(
    hour_id = floor(tau_i / (hz_val * 60 * 60)),
    hour_id = hour_id %% 24
  )

# aggregate the cadence measurements by computing empirical mode
# (1 value per day-hour)
out_comb_1_mode_hourly <- 
  out_comb_1 %>%
  group_by(loc_id, day_id, hour_id) %>%
  summarise(cadence_mode = estimate_mode(cadence_i),
            cnt = n()) %>%
  as.data.frame()

ggplot(out_comb_1_mode_hourly %>% filter(cnt >= 10), 
       aes(x = hour_id, y = cadence_mode, group = 1)) + 
  geom_line(linetype = 2, size = 0.5, color = "grey") + 
  geom_point(aes(size = cnt, alpha = cnt)) + 
  facet_grid(loc_id ~ day_id) + 
  theme_minimal(base_size = 10) + 
  labs(x = "Hour of a day", 
       y = "Cadence [steps/s]",
       alpha = "Segmented\nstrides\ncount",
       size = "Segmented\nstrides\ncount") + 
  scale_x_continuous(limits = c(0, 23), breaks = seq(0, 23, by = 2))
```


### Session info

<details>
  <summary>(Click to see session info.)</summary>

```{r}
devtools::session_info()
```
</details>


### References

[1]: Karas, M., Urbanek, J.K., Illiano, V.P., Bogaarts, G., Crainiceanu, C.M., Dorn, J.F. (2021). *Estimation of free-living walking cadence from wrist-worn sensor accelerometry data and its association with SF-36 quality of life scores*, Submitted. 

[2]: Karas, M., Straczkiewicz, M., Fadel, W., Harezlak, J., Crainiceanu, C.M., Urbanek, J.K.
(2018). Adaptive empirical pattern transformation (ADEPT) with application to walking stride
segmentation, Biostatistics, kxz033. 






